---
title: "Introduction to R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Programming

R is an open source programming language and software environment for statistical computing and graphics that is supported by the R Foundation for Statistical Computing. Today, it is one of the most popular languages, being used all across the world in a wide variety of domains and fields.

RStudio is an integrated development environment (IDE) for R. It includes a console, syntax-highlighting editor that supports direct code execution, as well as tools for plotting, history, debugging and workspace management.

## Working directory
```{r}
getwd()
```

## Calculating things in R
```{r}
3+3
```

```{r}
1/100
```

```{r}
sqrt(4)
```

## Variable and Data Types

There are several different types of data you can use in R. We'll examine a few common ones in a little more detail.

### Text
Strings are known as "character" in R. Use the double quotes `"` or single quotes `'` to wrap around the string
```{r}
myname <- "João"
```

We can use the `class()` function to see what data type it is
```{r}
class(myname)
```

### Numbers
Numbers have different classes. The most common two are `integer` and `numeric`. Integers are whole numbers:
```{r}
favourite.integer <- as.integer(8)
print(favourite.integer)
class(favourite.integer)
```

Numbers can be `numeric` which are decimals:
```{r}
favourite.numeric <- as.numeric(8.8)
print(favourite.numeric)
class(favourite.numeric)
```

### Logical (True/False)
We use the `==` to test for equality in R
```{r}
class(TRUE)
favourite.numeric == 8.8
favourite.numeric == 9.9
```

### Factors
In order to categorise the categorical variables and store it on multiple levels, we use the data object called R factor.

```{r}
directions <- as.factor(c("North", "North", "West", "South"))
class(directions)
```

How many categories are there for disorders and what are they?
```{r}
levels(directions)
nlevels(directions)
```

A factor can be ordered. This makes sense in the context of a ranking such as a survey response, e.g. from 'Strongly agree' to 'Strong disagree'.
```{r}
myorderedfactor <- factor(directions, levels = c("South", "West", "North"), ordered = TRUE)

levels(myorderedfactor)
```

### Vectors
We can create 1D data structures called "vectors".
```{r}
1:10
2*(1:10)
seq(0, 10, 2)
```
```{r}
directions <- c("North", "North", "West", "South")
class(directions)
```

### Matriz
A matrix in R is a two-dimensional rectangular data set and thus it can be created using vector input to the matrix function

```{r}
mat <- matrix (
c(2, 4, 3, 1, 5, 7),        # the data elements
nrow = 2,                   # no. of rows
ncol = 3,                   # no. of columns
byrow = TRUE)               # arranging it by row

print(mat)
```

How to Access Elements of Matrix in R?

```{r}
mat[2, 3]

mat[2, ]

mat[ ,3]
```

### Dataframe
A data frame is being used for storing data tables, the vectors that are contained in the form of a list in a data frame are of equal length.

#### Data Transformation

```{r}
#install.packages("tidyverse")
#install.packages("nycflights13")

library(tidyverse)
library(nycflights13)

summary(flights)

filter(flights, month == 1, day == 1)

select(flights, year:day)

select(flights, -c(year:day))

rename(flights, tail_num = tailnum)

flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)

mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)

transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)

delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
delays

```
