---
title: "Statistical Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Logistic Regression in R with Healthcare data
We will run a logistic regression to evaluate the effect of calcium and vitD on the osteoporosis. In this tutorial we will demonstrate some of the many options the **RNHANES** package. This tutorial was adapted from **Anisa Dhana's** tutorial at https://datascienceplus.com/.

## Prerequisites
```{r}
library(tidyverse)
library(RNHANES)
library(ggplot2)
library(pROC)
```

## Prepare the dataset

```{r}
d07 = nhanes_load_data("DEMO_E", "2007-2008") %>%
  select(SEQN, cycle, RIAGENDR, RIDAGEYR) %>%
  transmute(SEQN=SEQN, wave=cycle, RIAGENDR, RIDAGEYR) %>% 
  left_join(nhanes_load_data("VID_E", "2007-2008"), by="SEQN") %>%
  select(SEQN, wave, RIAGENDR, RIDAGEYR, LBXVIDMS) %>% 
  transmute(SEQN, wave, RIAGENDR, RIDAGEYR, vitD=LBXVIDMS) %>% 
  left_join(nhanes_load_data("BIOPRO_E", "2007-2008"), by="SEQN") %>%
  select(SEQN, wave, RIAGENDR, RIDAGEYR, vitD, LBXSCA) %>% 
  transmute(SEQN, wave, RIAGENDR, RIDAGEYR, vitD, Calcium = LBXSCA) %>% 
  left_join(nhanes_load_data("OSQ_E", "2007-2008"), by="SEQN") %>%
  select(SEQN, wave, RIAGENDR, RIDAGEYR, vitD, Calcium, OSQ060) %>% 
  transmute(SEQN, wave, RIAGENDR, RIDAGEYR, vitD, Calcium, Osteop = OSQ060)

d09 = nhanes_load_data("DEMO_F", "2009-2010") %>%
  select(SEQN, cycle, RIAGENDR, RIDAGEYR) %>%
  transmute(SEQN=SEQN, wave=cycle, RIAGENDR, RIDAGEYR) %>% 
  left_join(nhanes_load_data("VID_F", "2009-2010"), by="SEQN") %>%
  select(SEQN, wave, RIAGENDR, RIDAGEYR, LBXVIDMS) %>% 
  transmute(SEQN, wave, RIAGENDR, RIDAGEYR, vitD=LBXVIDMS) %>% 
  left_join(nhanes_load_data("BIOPRO_F", "2009-2010"), by="SEQN") %>%
  select(SEQN, wave, RIAGENDR, RIDAGEYR, vitD,  LBXSCA) %>% 
  transmute(SEQN, wave, RIAGENDR, RIDAGEYR, vitD, Calcium = LBXSCA) %>% 
  left_join(nhanes_load_data("OSQ_F", "2009-2010"), by="SEQN") %>%
  select(SEQN, wave, RIAGENDR, RIDAGEYR, vitD, Calcium, OSQ060) %>% 
  transmute(SEQN, wave, RIAGENDR, RIDAGEYR, vitD, Calcium, Osteop = OSQ060)

dat = bind_rows(d07, d09) %>% as.data.frame()
```

### Create categories of Vitamin D
```{r}
dat1 = dat %>% 
  mutate(
    vitD_group = case_when(
      vitD < 30 ~ "Deficiency",
      vitD >= 30 & vitD < 50 ~ "Inadequacy",
      vitD >= 50 & vitD <= 125 ~ "Sufficiency"))
```

### Exclude missings
```{r}
dat2 = dat1 %>% 
  filter(!is.na(vitD_group), !is.na(Calcium), !is.na(Osteop), Osteop!=9) %>% 
  mutate(Gender = recode_factor(RIAGENDR, 
                           `1` = "Men", 
                           `2` = "Women"),
         Osteop = recode_factor(Osteop, 
                           `1` = 1, 
                           `2` = 0))

head(dat2)
```

## Logit regression model
I will use the **glm()** function to run the logistic regression and then summary() command to get the results.

```{r}
fit <- glm(Osteop ~ vitD_group + Calcium + Gender + RIDAGEYR, 
           data = dat2, 
           family = "binomial")
summary(fit)

round(exp(coef(fit)), 2)
```

